### 2020-06-28 Case Studies/06_pairs-trading.rmd
case studies, pair trading / optimized simulations with C
case studies, pair trading / optimized simulations with C

@@

```{r}
getSymbols("T", src = "yahoo", from = "1985-01-01", to = "2015-12-31")
getSymbols("VZ", src = "yahoo", from = "1985-01-01", to = "2015-12-31")
```

```{r, fig.height=4}
chartSeries(T)
chartSeries(VZ)
```

@@

```{r}
combineStocks <- 
   function(lseries, rseries, 
            stockNames = c(deparse(substitute((a)),
                           deparse(substitute(b)))),
            add_ratio = T) {
   l_adj <- lseries[, 6]; r_adj <- rseries[, 6]
   combined <- merge(l_adj, r_adj)

@@

@@

@@

@@

      if(startDay > 1)
         ratio <- ratio[ - (1:(startDay-1))]

      isExtreme <- ratio >= up | ratio <= down

      if(!any(isExtreme))
         return(integer())

@@

@@

@@

showPosition(overlap$Date[a], r[a])
showPosition(overlap$Date[b], r[b])
showPosition(overlap$Date[c], r[c])
```

@@

```{r}
getPositions <-
   function(ratio, k = 1, m = mean(ratio), s = sd(ratio))
   {
      when = list()
      cur = 1

@@

```{r}
pos <- getPositions(r, k)
plotRatio(r, k, overlap$Date, xlab = "Date", ylab = "Ratio")
invisible(lapply(pos, function(p)
                           showPosition(overlap$Date[p], r[p])))
```

@@

```{r}
positionProfit <-
   # r = overlap$att / overlap$verizon
   # k = 1.7
   # pos = getPositions(r, k)
   # positionProfile(pos[[1]], overlap$att, overlap$verizon)
   function(pos, stockPriceA, stockPriceB,
            ratioMean = mean(stockPriceA / stockPriceB ),
            p = 0.001, byStock = F)
   {
      if(is.list(pos)) {
         ans = sapply(pos, positionProfit,
                      stockPriceA, stockPriceB, ratioMean, p, byStock)

@@

@@

>>>>

```{r read-data}

@@

```{r get-symbols}

getSymbols("T", src = "yahoo", from = "1985-01-01", to = "2015-12-31")
getSymbols("VZ", src = "yahoo", from = "1985-01-01", to = "2015-12-31")
```

```{r plot-symbols, fig.height=4}

chartSeries(T)
chartSeries(VZ)
```

@@

```{r combine-stocks}

combineStocks <- 
   function(lseries, rseries, 
            stockNames = c(deparse(substitute((a)),
                           deparse(substitute(b)))),
            add_ratio = T) {
   l_adj <- lseries[, 6]; r_adj <- rseries[, 6]
   combined <- merge(l_adj, r_adj)

@@

```{r get-overlap}

@@

```{r plot-ratio}

@@

```{r find-positions}

@@

      if(startDay > 1)
           ratio = ratio[ - (1:(startDay-1)) ]

     isExtreme = ratio >= up | ratio <= down

     if( is.na(isExtreme) || !any(isExtreme))
         return(integer())

@@

```{r show-position}

@@

```{r plot-ratio-2}

@@

#showPosition(overlap$Date[a], r[a])
#showPosition(overlap$Date[b], r[b])
#showPosition(overlap$Date[c], r[c])
```

@@

```{r get-positions}

getPositions <-
   function(ratio, k = 1, m = mean(ratio), s = sd(ratio))
   {
      when = list()
      cur = 1

@@

@@

positionProfit <-
   # r = overlap$att / overlap$verizon
   # k = 1.7
   # pos = getPositions(r, k)
   # positionProfile(pos[[1]], overlap$att, overlap$verizon)
   function(pos, stockPriceA, stockPriceB,
            ratioMean = mean(stockPriceA / stockPriceB ),
            p = 0.001, byStock = F)
   {
      if(is.list(pos)) {
         ans = sapply(pos, positionProfit,
                      stockPriceA, stockPriceB, ratioMean, p, byStock)

@@

      if(is.na(priceA[1]) | is.na(priceB[1]))
         return(0)

@@

## Simulation

Vector Auto-regression

$X_t^{(1)} = \rho X^{(1)}_{t-1} + \psi(1 - \rho)X^{(2)}_{t-1} + \epsilon^{(1)}_t$

$X_t^{(2)} = \rho X^{(2)}_{t-1} + \psi(1 - \rho)X^{(1)}_{t-1}+ \epsilon^{(2)}_t$

$\epsilon^{(1)}_t \sim N(0, \sigma^2_i)$

...

$Y^{(1)}_t = \beta^{(1)}_0 + \beta^{(1)}_1t + X^{1}_t$

$Y^{(2)}_t = \beta^{(2)}_0 + \beta^{(2)}_1t + X^{2}_t$

### Simulating the Stock Price Series

```{r}
stockSim <-
   function(n = 4000, rho = 0.99, psi = 0, sigma = rep(1, 2),
            beta0 = rep(100, 2), beta1 = rep(0, 2),
            epsilon = matrix(rnorm(2*n, sd = sigma),
                             nrow = n, byrow = T))
   {
      X <- matrix(0, nrow = n, ncol = 2)
      X[1, ] <- epsilon[1, ]

      A <- matrix(c(rho, psi * (1-rho), psi*(1-rho), rho), nrow = 2)
      for(i in 2:n)
         X[i, ] = A %*% X[i-1, ] + epsilon[i, ]

      X[, 1] <- beta0[1] + beta1[1] * (1:n) + X[, 1]
      X[, 2] <- beta0[2] + beta1[2] * (1:n) + X[, 2]

      X
   }

```{r}
a <- stockSim(rho = .99, psi = 0)

matplot(1:nrow(a), a, type = "l", xlab = "Day", ylab = "Y",
        col = c("black", "grey"), lty = "solid")

beta1 <- c(.05, .1)

a <- stockSim(beta1 = c(.05, .1))

matplot(1:nrow(a), a, type = "l", xlab = "Day", ylab = "Y",
        col = c("black", "grey"), lty = "solid")

### Simulation Utilities

runSim <-
function(rho = .99, psi = .9, beta0 = c(100, 100), beta1 = c(0, 0),
         sigma = c(1, 1), n = 4000)
{
    X = stockSim(n, rho, psi, sigma, beta = beta0, beta1 = beta1)
    train = X[ 1:floor(n/2), ]
    test = X[ (floor(n/2)+1):n, ]
    m = mean(train[, 1]/train[, 2])
    s = sd(train[, 1]/train[, 2])
    k.star = getBestK(train[, 1], train[, 2], m = m, s = s)
    getProfit.K(k.star, test[, 1], test[, 2], m, s)
}

getProfit.K =
function(k, x, y, m = mean(x/y), s = sd(x/y))
{
    pos = getPositions(x/y, k, m = m, s = s)
    if(length(pos) == 0)  
       0
    else
       sum(positionProfit(pos, x, y, m), na.rm=T)
}

getBestK = 
function(x, y, ks = seq(0.1, max.k, length = N), N = 100, 
         max.k = NA, m = mean(x/y), s = sd(x/y))
{
    if(is.na(max.k)) {
       r = x/y
       max.k = max(r/sd(r))
    }

    pr.k = sapply(ks, getProfit.K, x, y, m = m, s = s)
    median(ks[ pr.k == max(pr.k) ])
}

simProfitDist = 
function(..., B = 999) 
      sapply(1:B,  function(i, ...) runSim(...), ...)

### Run the Simulation

```{r}
system.time({ x = simProfitDist( .99, .9, c(0, 0)) })

ggplot(data.table(profit = x), aes(profit)) +
   geom_density()
```

```{r}
g <- expand.grid(psi = seq(.8, .99, length.out = 20),
                 beta1 = seq(-0.01, .01, length.out = 20),
                 beta2 = seq(-0.1, 0.01, length.out = 20))

dim(g)
```

```{r}
Rprof("sim.prof")

system.time({x = simProfitDist( .99, .9, c(0, 0))})

Rprof(NULL)

head(summaryRprof("sim.prof")$by.self)
```

```{r}
counter <- 0L
trace(findNextPosition, quote( counter <<- counter + 1L),
      print = FALSE)

system.time({x = simProfitDist( .99, .9, c(0, 0))})

counter

untrace(findNextPosition)
```

```{r}
library(compiler)
stockSim.cmp <- cmpfun(stockSim)

tm.orig <- system.time({replicate(80, stockSim())})
tm.compiled <- system.time({replicate(80, stockSim.cmp())})

tm.orig/tm.compiled
```

```{r}
c.lib <- paste(here(), "Case Studies", "06_stockSim.dll", sep="/")

dyn.load(c.lib)
```

```{r}
stockSim.c <-
   function(n = 4000, rho = 0.99, psi = 0, sigma = rep(1, 2),
            beta0 = rep(100, 2), beta1 = rep(0, 2),
            epsilon = matrix(rnorm(2*n, sd = sigma), nrow = n))
   {
      X <- matrix(0, nrow = n, ncol = 2)
      X[1, ] <- epsilon[1, ]

      X <- .C("stockSim", X, as.integer(n), rho, psi, epsilon)[[1]]

      X[, 1] <- beta0[1] + beta1[1] + (1:n) + X[, 1]
      X[, 2] <- beta0[2] + beta1[2] + (1:n) + X[, 2]

      X
   }
```

e <- matrix(rnorm(2*4000, sd = c(1, 1)), , 2)
tmp1 <- stockSim.c(epsilon = e)
tmp2 <- stockSim(epsilon = e)

stockSim <- stockSim.c

Rprof("sim.prof")
system.time({x = simProfitDist(.99, .9, c(0, 0))})
Rprof(NULL)
head(summaryRprof("sim.prof")$by.self)
```
